**Principal Component Analysis (PCA) and K-Nearest Neighbors (KNN) Classifier**

Principal Component Analysis (PCA) is a dimensionality reduction technique used to transform high-dimensional data into a lower-dimensional space while preserving most of the important information. It achieves this by identifying the directions (principal components) that capture the maximum variance in the data. PCA is widely employed for exploratory data analysis and visualization, making it easier to comprehend complex datasets.

K-Nearest Neighbors (KNN) is a simple yet effective classification algorithm used for both classification and regression tasks. In KNN, the class of a data point is determined by the majority class among its k-nearest neighbors in the feature space. This non-parametric algorithm is particularly useful when the decision boundary is irregular or the data distribution is not well-defined. KNN's simplicity and intuitive nature make it a popular choice for various machine learning applications.

Importing Libraries

```{r}
# Loading necessary libraries
library(corrr)      # For correlation analysis
library(ggcorrplot) # For visualizing correlations
library(factoextra) # For PCA visualization
library(class)      # For KNN classifier
library(ISLR2)      # For Smarket dataset
library(caret)      # For machine learning workflow
library(MASS)       # For logistic regression
library(caTools)    # For data splitting
# Loading Smarket dataset
data("Smarket")
attach(Smarket)  # Attaching the dataset to access variables directly
```

PCA on SMarket Dataset

```{r}
# Preparing data for PCA
Smarket_pca = Smarket[, -9]  # Excluding the 'Year' column
Smarket_pca_scaled = scale(Smarket_pca)  # Scaling the data
Smarket_pca_results = princomp(Smarket_pca_scaled)  # Performing PCA

summary(Smarket_pca_results)  # Summary of PCA results

# Plotting PCA results
plot(Smarket_pca_results, type = "l")  # Scree plot
biplot(Smarket_pca_results, scale = 0)  # Biplot
fviz_pca_var(Smarket_pca_results, col.var = "black")  # Variable factor map
fviz_cos2(Smarket_pca_results, choice = "var", axes = 1:2)  # Contributions of variables to dimensions
fviz_pca_var(Smarket_pca_results, col.var = "cos2",
             gradient.cols = c("black", "orange", "green"),
             repel = TRUE)  # Quality of representation of variables
```

KNN Classifier on Smarket

```{r}
# Splitting data into training and testing sets
train = Smarket[Smarket$Year < 2005, ]
test = Smarket[Smarket$Year == 2005, ]

k = 3  # Setting the value of k

# Performing KNN classification
knn.fit = knn(train[,-9], test[,-9], cl = train$Direction, k = k)

# Calculating training and testing errors
train.pred = knn(train[,-9], train[,-9], cl = train$Direction, k = k)
train.error = mean(train.pred != train$Direction)
test.error = mean(knn.fit != test$Direction)

cat("Training error:", train.error, "\n")
cat("Testing error:", test.error, "\n")
```

**Logistic Regression Model**: Logistic regression is a statistical method used for binary classification tasks. Unlike linear regression, which predicts continuous values, logistic regression predicts the probability of occurrence of an event by fitting data to a logistic curve. It's widely employed in fields such as medicine, social sciences, and economics due to its simplicity and interpretability. In this context, logistic regression is applied to predict the direction of stock market movements based on lagged stock returns and trading volume.

Logistic Model on Smarket Dataset

```{r}
# Fitting logistic regression model
glm.fit = glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = train, family = binomial)

# Predicting on training and testing sets
train.pred = predict(glm.fit, type = "response", newdata = train)
test.pred = predict(glm.fit, type = "response", newdata = test)

# Calculating training and testing errors
train.errors = mean((train$Direction != ifelse(train.pred > 0.5, "Up", "Down")))
test.errors = mean((test$Direction != ifelse(test.pred > 0.5, "Up", "Down")))

cat("Training error: ", train.errors, "\n")
cat("Testing error: ", test.errors, "\n")
```
